Lesson 6
========================================================
```{r Packages}
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(dplyr)
library(GGally)
library(reshape2) ## Package to transform data between wide and long formats
library(scales)
library(memisc)
library(RColorBrewer)
library(bitops)
library(RCurl)
```

***

### Scatterplot Review

```{r Scatterplot Review}
# Let's start by examining two variables in the data set.
# The scatterplot is a powerful tool to help you understand
# the relationship between two continuous variables.

# We can quickly see if the relationship is linear or not.
# In this case, we can use a variety of diamond
# characteristics to help us figure out whether
# the price advertised for any given diamond is 
# reasonable or a rip-off.

# Let's consider the price of a diamond and it's carat weight.
# Create a scatterplot of price (y) vs carat weight (x).

# Limit the x-axis and y-axis to omit the top 1% of values.

# ENTER YOUR CODE BELOW THIS LINE
# ================================================================

ggplot(aes(x = carat, y = price), data = diamonds) +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99)) +
  geom_point(fill = I('#F79423'), color = 'black', shape = 21, alpha = 1/4) + 
  stat_smooth(method = 'lm')

ggplot(aes(x = carat, fill = cut), data = diamonds) + 
  geom_density() + 
  scale_x_log10()

```

***************************************************************************************************

### Price and Carat Relationship
Response:

Del grafico anterior puede verse como la relacional no es lineal (quizas sea exponencial).
The dispersion or variance of the relationship also increases as carat size increases.

***************************************************************************************************

### ggpairs Function
Notes:

```{r ggpairs Function}
# install these if necessary
# install.packages('GGally')
# install.packages('scales')
# install.packages('memisc')
# install.packages('lattice')
# install.packages('MASS')
# install.packages('car')
# install.packages('reshape')
# install.packages('plyr')


# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, params = c(shape = I('.'), outlier.shape = I('.')))
```


***************************************************************************************************

### The Demand of Diamonds
Notes:

```{r The Demand of Diamonds}
# Create two histograms of the price variable
# and place them side by side on one output image.

# We've put some code below to get you started.

# The first plot should be a histogram of price
# and the second plot should transform
# the price variable using log10.

# Set appropriate bin widths for each plot.
# ggtitle() will add a title to each histogram.

# You can self-assess your work with the plots
# in the solution video.

# ALTER THE CODE BELOW THIS LINE
# ==============================================

p1 = ggplot(aes(x = price), data = diamonds) + 
      geom_histogram(color = 'black', fill = I('#F952B3'), binwidth = 150) + 
      ggtitle('Price')

p2 = ggplot(aes(x = price), data = diamonds) + 
      geom_histogram(color = 'black', fill = I('#F952B3'), binwidth = 0.02) +
      scale_x_log10() +
      ggtitle('log10(Price)')

grid.arrange(p1, p2, ncol = 1)


```

***************************************************************************************************

### Connecting Demand and Price Distributions
Notes:
- Sin aplicar la transformacion de log10, se aprecia como tenemos una distribucion 'skewed' (con mucha cola). Entonces aplicamos la tranformacion logaritmica de los datos.

- Con la transformacion logaritmica se aprecia como los datos se acerca a una distribución normal (normal distribution) con forma de campana, incluso podríamos decir que hay dos campanas (bimodal). Al comprar diamantes, eso podría representar que los clientes estan en dos clases economicas diferentes.

***************************************************************************************************

### Scatterplot Transformation

```{r Scatterplot Transformation}
ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point() +
  scale_y_log10() +
  ggtitle('log10(Price) by Carat')
```

Para graficar en base el volumen, hay que tener que el volumen está en escala cubica. Por lo que necesitamos una funcion para transformar la variable 'carat':

### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() {
  trans_new('cuberoot', transform = function(x) x^(1/3),
            inverse = function(x) x^3 )
}
# Esta funcion calcula los valores de X al cubo y su inversa para poder graficarlo bien.
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')

#Otra forma:
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_log10(limits = c(350, 15000),
                breaks = c(350, 1000, 5000, 10000, 15000)) +
  geom_smooth(method = 'lm') +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***************************************************************************************************

### Overplotting Revisited
Overplotting: 
When multiple points take on the same values. This is often due to rounding.

```{r Sort and Head Tables}
# table: resumir en una tabla cuantas veces se repite cierto valor de una variable
# sort: ordenar la tabla
# head: mostrar los 6 primeras valores (por defecto)

# Mostramos los 6 valores de las variables carat y price mas repetidos:
head(sort(table(diamonds$carat), decreasing = TRUE) )
head(sort(table(diamonds$price), decreasing = TRUE) )

# Resultado:
#  0.3 0.31 1.01  0.7 0.32    1  --> valor de la variable carat
# 2604 2249 2242 1981 1840 1558  --> las veces que se repite en el data set
```

Como se pude apreciar, hay valores que estan demasiado repetidos, lo que hace que el resultado sea overplotting. 

El hecho de que haya overplotting puede "eclipsar" en cierto modo la densidad y dispersion de nuestros datos. 

Como se ha visto en lecciones anteriores, se puede utilizar la funcionalidad de jitter (para introducir una especie de ruido entre los puntos) o añadiendo transparencia (que se necesiten por ejemplo 4 puntos para que se vea del todo --> parametro alpha en ggplot).


```{r Overplotting Revisited}
# Add a layer to adjust the features of the
# scatterplot. Set the transparency to one half,
# the size to three-fourths, and jitter the points.

# If you need hints, see the Instructor Notes.
# There are three hints so scroll down slowly if
# you don't want all the hints at once.

# ALTER THE CODE BELOW THIS LINE
# =======================================================================
ggplot(aes(carat, price), data = diamonds) + 
  geom_jitter(alpha = 1/2) +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_log10(limits = c(350, 15000),
                breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')


# Otra forma:
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 1/2, position = position_jitter(h = 0.02), size = 3/4) +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_log10(limits = c(350, 15000),
                breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')

# Otra forma:
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 1/2, position = 'jitter', size = 3/4) +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_log10(limits = c(350, 15000),
                breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***************************************************************************************************

### Examinar el precio de los diamantes segun los quilates y su claridad:
```{r}
ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')


# Con la paleta de colores Brewer:
ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', 
                                          reverse = TRUE,
                                          override.aes = list(alpha = 1, size = 4) ) ) +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```


***************************************************************************************************

### Examinar el precio de los diamantes segun los quilates y su corte:

Alter the code below.
```{r Price vs. Carat and Cut}
# Let's look at cut and see if we find a similar result.

# Adjust the code below to color the points by cut.
# Change any other parts of the code as needed.

# ALTER THE CODE BELOW THIS LINE
# =========================================================================================
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', 
                                          reverse = TRUE,
                                          override.aes = list(alpha = 1, size = 4) ) ) +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')

```


***************************************************************************************************

### Examinar el precio de los diamantes segun los quilates y su Color:

Alter the code below.
```{r Price vs. Carat and Color}
# Finally, let's use diamond color to color our plot.

# Adjust the code below to color the points by diamond colors
# and change the titles.

# ALTER THE CODE BELOW THIS LINE
# ===========================================================================================
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', 
                                          reverse = FALSE,
                                          override.aes = list(alpha = 1, size = 4) ) ) +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')

```


***************************************************************************************************

### Linear Models in R
Notes:
- Función lm( y ~ x ). y = outcome variable // x = explanatory variable.


Response:
Which of these formulas would we use inside the lm() functio?
log(price) ~ carat^(1/3)

***************************************************************************************************

### Building the Linear Model
Notes:

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)

# Añadir nuevas variables al modelo:
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)

# Function I:
# In function formula. There it is used to inhibit the interpretation of operators such 
# as "+", "-", "*" and "^" as formula operators, so they are used as arithmetical operators.

summary(m1, correlation = TRUE)
summary(m2, correlation = TRUE)
summary(m3, correlation = TRUE)
summary(m4, correlation = TRUE)
summary(m5, correlation = TRUE)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***************************************************************************************************

### Model Problems
Video Notes:
Let's put our model in a larger context. Assuming that the data is not somehow corrupted and we are not egregiously violating some of the key assumptions of linear regression (for example, violating the IID assumption by having a bunch of duplicated observations in our data set), what could be some problems with this model? What else should we think about when using this model? 

El parametro "Intercept" no tiene ninguna logica en el modelo en caso de que los parametros de la variable explicatoria del modelo (X) no puedan tener el valor 0, o no tengan el valor 0 en ningun caso del data set.

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:

***************************************************************************************************

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}
# install.package('bitops')
# install.packages('RCurl')

diamondsurl = getBinaryURL('https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda', ssl.verifypeer = FALSE)

load(rawConnection(diamondsurl))
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}
# Your task is to build five linear models like Solomon
# did for the diamonds data set only this
# time you'll use a sample of diamonds from the
# diamondsbig data set.

# Be sure to make use of the same variables
# (logprice, carat, etc.) and model
# names (m1, m2, m3, m4, m5).

# To get the diamondsbig data into RStudio
# on your machine, copy, paste, and run the
# code in the Instructor Notes. There's
# 598,024 diamonds in this data set!

# Since the data set is so large,
# you are going to use a sample of the
# data set to compute the models. You can use
# the entire data set on your machine which
# will produce slightly different coefficients
# and statistics for the models.

# This exercise WILL BE automatically graded.

# You can leave off the code to load in the data.
# We've sampled the data for you.
# You also don't need code to create the table output of the models.
# We'll do that for you and check your model summaries (R^2 values, AIC, etc.)

# Your task is to write the code to create the models.

# DO NOT ALTER THE CODE BELOW THIS LINE (Reads in a sample of the diamondsbig data set)
#===========================================================================================
setwd('C:/WORKSPACE/RSTUDIO/EDA_Course_Materials/lesson6')
diamondsbig <- read.csv('diamondsbig.csv')

diamondsbig$logprice = log(diamondsbig$price)

m1 <- lm( logprice ~ I(carat^(1/3)), 
          data = diamondsbig[diamondsbig$price < 1000 &
                             diamondsbig$cert == "GIA", ] )

# Añadir nuevas variables al modelo:
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)

# DO NOT ALTER THE CODE BELOW THIS LINE (Tables your models and pulls out the statistics)
#===========================================================================================
suppressMessages(library(lattice))
suppressMessages(library(MASS))
suppressMessages(library(memisc))
models <- mtable(m1, m2, m3, m4, m5)

```


***************************************************************************************************

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you ha€™ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
thisDiamond2 = data.frame(carat = 5.00, cut = "Good",
                         color = "I", clarity="VS1")

modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
modelEstimate2 = predict(m5, newdata = thisDiamond2,
                        interval="prediction", level = .95)
print(modelEstimate)
print(modelEstimate2)

# The prediction interval here may be slightly conservative, as the model errors are heteroskedastic over carat (and hence price) even after our log and cube-root transformations. 


dat = data.frame(m4$model, m4$residuals)
with(dat, sd(m4$residuals))
with(subset(dat, carat > .9 & carat < 1.1), sd(m4$residuals))


ggplot(aes(y = price, x = round(carat, 2)), data = diamondsbig) + 
  geom_line(stat = "summary", fun.y = sd)

```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI (confidence interval).


***************************************************************************************************

## APUNTES
===========
Funciones en R:
---------------
https://www.youtube.com/watch?v=DzIy6U-N6ac&list=PLOU2XLYxmsIK9qQfztXeybpHvru-TrqAP

Transformaciones:
-----------------
http://blog.ggplot2.org/post/25938265813/defining-a-new-transformation-for-ggplot2-scales

Colores en ggplot2: [Paletas de colores Brewer]
-------------------
http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/#palettes-color-brewer
http://docs.ggplot2.org/current/scale_brewer.html

Linear Models in R:
-------------------
http://data.princeton.edu/R/linearModels.html

Interpretar los coeficientes de Linear Regresion:
-------------------------------------------------
http://www.theanalysisfactor.com/interpreting-regression-coefficients/
http://blog.yhathq.com/posts/r-lm-summary.html
http://stats.stackexchange.com/questions/24242/how-to-apply-coefficient-term-for-factors-and-interactive-terms-in-a-linear-equa/24256#24256
